### 2020-03-05 Experiment_2/pool_ig_ms_results.Rmd

---
title: "comparing models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### fit MS models using DEoptim and compre to IG models using optim  
```{r}

pool_de_optim <- data.frame()
for(id in unique(dat_exp2$human_id)) {
pool_de_optim <- rbind(pool_de_optim,
                       data.frame(
                         "ID" = id,
                         "condition" = dat_exp2[dat_exp2$human_id==id,"condition"][1],
                         "Random_BIC" = -2*(180*log(1/3)),

                         # Bayesian updating with between and within game transfer
                         "Btwn_TR_BIC" = exp2_BT[[id]]$value + 2*log(180),
                         "Distinct_game_BIC" = Bayes_distinct_game[[id]]$optim$bestval + 2*log(180),

                         # # Bayesian updating with within but no btwn game transfer
                         "No_Btwn_Tr_BIC" = exp2_NO_BT[[id]]$value+ 2*log(180),
                         "Distinct_stage_BIC" = Bayes_distinct_stage[[id]]$optim$bestval + 2*log(180),
                         # 
                         # # No within or between game transfer
                         "No_Tr_BIC" = exp2_NO_Tr[[id]]$value+ 2*log(180),
                         "Bayes_distinct_no" = Bayes_distinct_no[[id]]$optim$bestval + 2*log(180),
                         # 
                         # # No distinction between players, transfer within and between
                         "Naive_BIC" = exp2_Naive[[id]]$value+ 2*log(180),
                         "Bayes_same_game" = Bayes_same_game[[id]]$optim$bestval + 2*log(180)
                       ))
}

write.csv(pool_de_optim,file="pool_de_optim.csv",row.names = FALSE)

```

# Fit both MS and IG models using optim 
```{r}
pool_optim <- data.frame()
for(id in unique(dat_exp2$human_id)) {
pool_optim <- rbind(pool_optim,
                       data.frame(
                         "ID" = id,
                         "condition" = dat_exp2[dat_exp2$human_id==id,"condition"][1],
                         "Random_BIC" = -2*(180*log(1/3)),

                         # Bayesian updating with between and within game transfer
                         "Btwn_TR_BIC" = exp2_BT[[id]]$value + 2*log(180),
                         "Distinct_game_BIC" = Bayes_distinct_game[[id]]$value + 2*log(180),

                         # # Bayesian updating with within but no btwn game transfer
                         "No_Btwn_Tr_BIC" = exp2_NO_BT[[id]]$value+ 2*log(180),
                         "Distinct_stage_BIC" = Bayes_distinct_stage[[id]]$value + 2*log(180),
                         # 
                         # # No within or between game transfer
                         "No_Tr_BIC" = exp2_NO_Tr[[id]]$value+ 2*log(180),
                         "Bayes_distinct_no" = Bayes_distinct_no[[id]]$value + 2*log(180),
                         # 
                         # # No distinction between players, transfer within and between
                         "Naive_BIC" = exp2_Naive[[id]]$value+ 2*log(180),
                         "Bayes_same_game" = Bayes_same_game[[id]]$value + 2*log(180)
                       ))
}

write.csv(pool_optim,file="pool_optim.csv",row.names = FALSE)
```