### 2020-08-13 exp1_lik_by_trial.Rmd
added plots lik by trial

---
title: "exp1_lik_by_trial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
```

```{r}
# Get Data for experiment 1, remove unecessary columns

dat <- read.csv("exp1_data.csv")
#####
# dat <- subset(dat, human_id == "38VxtUSv_h6RR5-tAAA2")
####

optim_results <- read.csv("All_results.csv")

```

```{r}

#preparing data for functions to claculate likelihood

# dat <- as_tibble(dat) %>% group_by(human_id,game)
dat <- dat %>%
  mutate(ai_action_prev = lag(ai_action,1), human_action_prev = lag(human_action,1))

## level 0 predictions
dat <- dat %>% mutate(pred_a1_level0 = case_when(
  game == "rps" & ai_action_prev == "rock" ~ 1,
  game == "fwg" & ai_action_prev == "fire" ~ 1,
  game == "numbers" & ai_action_prev == "one" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a2_level0 = case_when(
  game == "rps" & ai_action_prev == "paper" ~ 1,
  game == "fwg" & ai_action_prev == "water" ~ 1,
  game == "numbers" & ai_action_prev == "two" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a3_level0 = case_when(
  game == "rps" & ai_action_prev == "scissors" ~ 1,
  game == "fwg" & ai_action_prev == "grass" ~ 1,
  game == "numbers" & ai_action_prev == "three" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a4_level0 = case_when(
  game == "numbers" & ai_action_prev == "four" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  TRUE ~ 0))
dat <- dat %>% mutate(pred_a5_level0 = case_when(
  game == "numbers" & ai_action_prev == "five" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  TRUE ~ 0))

## level 1 predictions
dat <- dat %>% mutate(pred_a1_level1 = case_when(
  game == "rps" & human_action_prev == "scissors" ~ 1,
  game == "fwg" & human_action_prev == "grass" ~ 1,
  game == "numbers" & human_action_prev == "five" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a2_level1 = case_when(
  game == "rps" & human_action_prev == "rock" ~ 1,
  game == "fwg" & human_action_prev == "fire" ~ 1,
  game == "numbers" & human_action_prev == "one" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a3_level1 = case_when(
  game == "rps" & human_action_prev == "paper" ~ 1,
  game == "fwg" & human_action_prev == "water" ~ 1,
  game == "numbers" & ai_action_prev == "two" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a4_level1 = case_when(
  game == "numbers" & ai_action_prev == "three" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  TRUE ~ 0))
dat <- dat %>% mutate(pred_a5_level1 = case_when(
  game == "numbers" & ai_action_prev == "four" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  TRUE ~ 0))

## level 2 predictions
dat <- dat %>% mutate(pred_a1_level2 = case_when(
  game == "rps" & ai_action_prev == "paper" ~ 1,
  game == "fwg" & ai_action_prev == "water" ~ 1,
  game == "numbers" & ai_action_prev == "four" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a2_level2 = case_when(
  game == "rps" & ai_action_prev == "scissors" ~ 1,
  game == "fwg" & ai_action_prev == "grass" ~ 1,
  game == "numbers" & ai_action_prev == "five" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a3_level2 = case_when(
  game == "rps" & ai_action_prev == "rock" ~ 1,
  game == "fwg" & ai_action_prev == "fire" ~ 1,
  game == "numbers" & ai_action_prev == "one" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a4_level2 = case_when(
  game == "numbers" & ai_action_prev == "two" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  TRUE ~ 0))
dat <- dat %>% mutate(pred_a5_level2 = case_when(
  game == "numbers" & ai_action_prev == "three" ~ 1,
  game == "numbers" & is.na(ai_action_prev) ~ 1/5,
  TRUE ~ 0))

dat$ai_action_num <- recode(dat$ai_action,"rock" = 1, "paper" = 2, "scissors" = 3, "fire" = 1, "water" = 2, "grass" = 3, "one" = 1, "two" = 2, "three" = 3, "four" = 4 , "five" = 5)
dat$human_action_num <- recode(dat$human_action,"rock" = 1, "paper" = 2, "scissors" = 3, "fire" = 1, "water" = 2, "grass" = 3, "one" = 1, "two" = 2, "three" = 3, "four" = 4 , "five" = 5)

```

```{r}
# All_results <- data.frame()
# for(id in unique(dat$human_id)) {
#   All_results <- rbind(All_results,
#                        data.frame(
#                          "ID" = id,
#                          "condition" = dat[dat$human_id==id,"condition"][1],
#                          "Random_BIC" = -2*(100*log(1/3) + 50*log(1/5)),

lik_by_trial <- data.frame()

for(id in unique(dat$human_id)) {

  tdat <- dat %>% subset(human_id == id)

  # QLS likelihoods 
  pars_QLS <- unlist(filter(optim_results, ID == id)%>%select(QL_states_alpha,QL_states_Beta))
  tdat$QLS_lik <- Q_learn_states(pars_QLS,tdat,"likelihood_by_trial",gamma = 0)

  # Bayes model no transfer likelihoods
  theta_NT <- unlist(filter(optim_results, ID == id)%>%select(theta_no_transfer))
  tdat$Bayes_NT_lik <- exp1_Bayes_model_LL(theta_NT ,data = tdat, generalize = "no", "likelihood_by_trial")

  # Bayes model with Transfer likelihood
  theta_Tr <- unlist(filter(optim_results, ID == id)%>%select(theta_transfer))
  tdat$Bayes_Tr_lik <- exp1_Bayes_model_LL(theta_Tr ,data = tdat, generalize = "game", "likelihood_by_trial")

  lik_by_trial <- rbind(lik_by_trial,tdat)
}
```

```{r}
library(ggplot2)

mean_lik <- lik_by_trial %>%
  select(human_id,condition,game, round,QLS_lik,Bayes_NT_lik,Bayes_Tr_lik) %>%
  group_by(condition,game,round) %>%
  summarise(mean_QLS_lik = mean(QLS_lik, na.rm = TRUE),
            mean_NT_lik = mean(Bayes_NT_lik, na.rm = TRUE),
            mean_Tr_lik = mean(Bayes_Tr_lik, na.rm = TRUE))

mean_lik$game <- factor(mean_lik$game,levels=c("rps","fwg","numbers"))

data_long <- gather(mean_lik, strategy, probability, mean_QLS_lik:mean_Tr_lik, factor_key=TRUE)

# make sure the different games are ordered in the way they were played
data_long$game <- factor(data_long$game,levels=c("rps","fwg","numbers"))
ggplot(data_long,aes(x=round,y=probability,colour=strategy)) + geom_line() + facet_grid(game~condition)

```