### 2020-09-04 Experiment_2/exp2_lik_by_trial.Rmd
fixed Q learning bugs

---
title: "exp2_lik_by_trial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)

dat <- read.csv("data_exp2.csv")

exp2_optim_results <- read.csv("exp2_results.csv")

```

```{r}

# Preparing data for running functions to calculate likelihoods

dat <- as_tibble(dat) %>% group_by(human_id,game)
dat <- dat %>%
   mutate(ai_action_prev = lag(ai_action,1), human_action_prev = lag(human_action,1))

## level 0 predictions
dat <- dat %>% mutate(pred_a1_level0 = case_when(
  game == "rps" & ai_action_prev == "rock" ~ 1,
  game == "fwg" & ai_action_prev == "fire" ~ 1,
  game == "shootout" & ai_action_prev == "left" ~ 1,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a2_level0 = case_when(
  game == "rps" & ai_action_prev == "paper" ~ 1,
  game == "fwg" & ai_action_prev == "water" ~ 1,
  game == "shootout" & ai_action_prev == "center" ~ 1,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a3_level0 = case_when(
  game == "rps" & ai_action_prev == "scissors" ~ 1,
  game == "fwg" & ai_action_prev == "grass" ~ 1,
  game == "shootout" & ai_action_prev == "right" ~ 1,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))

## level 1 predictions
dat <- dat %>% mutate(pred_a1_level1 = case_when(
  game == "rps" & human_action_prev == "scissors" ~ 1,
  game == "fwg" & human_action_prev == "grass" ~ 1,
  game == "shootout" & human_action_prev == "left" ~ 1,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a2_level1 = case_when(
  game == "rps" & human_action_prev == "rock" ~ 1,
  game == "fwg" & human_action_prev == "fire" ~ 1,
  game == "shootout" & human_action_prev == "center" ~ 1,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a3_level1 = case_when(
  game == "rps" & human_action_prev == "paper" ~ 1,
  game == "fwg" & human_action_prev == "water" ~ 1,
  game == "shootout" & ai_action_prev == "right" ~ 1,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))

## level 2 predictions
dat <- dat %>% mutate(pred_a1_level2 = case_when(
  game == "rps" & ai_action_prev == "paper" ~ 1,
  game == "fwg" & ai_action_prev == "water" ~ 1,
  game == "shootout" & (ai_action_prev == "center" | ai_action_prev == "right") ~ .5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a2_level2 = case_when(
  game == "rps" & ai_action_prev == "scissors" ~ 1,
  game == "fwg" & ai_action_prev == "grass" ~ 1,
  game == "shootout" & (ai_action_prev == "left" | ai_action_prev == "right") ~ .5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))
dat <- dat %>% mutate(pred_a3_level2 = case_when(
  game == "rps" & ai_action_prev == "rock" ~ 1,
  game == "fwg" & ai_action_prev == "fire" ~ 1,
  game == "shootout" & (ai_action_prev == "left" | ai_action_prev == "center") ~ .5,
  is.na(ai_action_prev) ~ 1/3,
  !is.na(ai_action_prev) ~ 0))

dat$ai_action_num <- recode(dat$ai_action,"rock" = 1, "paper" = 2, "scissors" = 3, "fire" = 1, "water" = 2, "grass" = 3, "left" = 1, "center" = 2, "right" = 3)
dat$human_action_num <- recode(dat$human_action,"rock" = 1, "paper" = 2, "scissors" = 3, "fire" = 1, "water" = 2, "grass" = 3, "left" = 1, "center" = 2, "right" = 3)

```

```{r}

dat <- as.data.frame(dat)   
lik_by_trial_2 <- data.frame()

for(id in unique(dat$human_id)) {

  tdat <- dat %>% subset(human_id == id)

  # QLS likelihoods 
  pars_QLS <- unlist(filter(exp2_optim_results, ID == id)%>%select(QL_states_Beta, QL_states_alpha))
  tdat$QLS_lik <-  exp2_QL_states(pars_QLS,tdat,"likelihood_by_trial",0.0, FALSE)

  # QLS within Transfer likelihoods 
  pars_QLS_within <- unlist(filter(exp2_optim_results, ID == id)%>%select(QLS_within_Tr_Beta,QLS_within_Tr_alpha))
  tdat$QLS_lik_within <-  exp2_QL_states(pars_QLS_within,tdat,"likelihood_by_trial",gamma = 0, TRUE)
  #                
  # 
  # Bayes model no transfer likelihoods
  theta_NT <- unlist(filter(exp2_optim_results, ID == id)%>%select(theta_no_tr))
  tdat$Bayes_NT_lik <- Bayes_model_LL(theta_NT, tdat, distinct_opponent = TRUE, generalize = "no",return = "likelihood_by_trial")

  # Bayes model with Transfer likelihood
  theta_Tr <- unlist(filter(exp2_optim_results, ID == id)%>%select(theta_transfer))
  tdat$Bayes_Tr_lik <- Bayes_model_LL(theta_Tr, tdat, distinct_opponent = TRUE, generalize = "game",return = "likelihood_by_trial")

  lik_by_trial_2 <- rbind(lik_by_trial_2,tdat)
}
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

mean_lik_2 <- lik_by_trial_2 %>%
  select(human_id,game,stage,round_condition, round, QLS_lik, QLS_lik_within, Bayes_NT_lik, Bayes_Tr_lik) %>%
  group_by(game,stage,round) %>%
  summarise(mean_QLS_lik = mean(QLS_lik, na.rm = TRUE),
            mean_QLS_within_lik = mean(QLS_lik_within, na.rm = TRUE),
            mean_NT_lik = mean(Bayes_NT_lik, na.rm = TRUE),
            mean_Tr_lik = mean(Bayes_Tr_lik, na.rm = TRUE))

mean_lik$game <- factor(mean_lik$game,levels=c("rps","fwg","shootout"))

data_long <- gather(mean_lik_2, strategy, probability, mean_QLS_lik:mean_Tr_lik, factor_key=TRUE)

# make sure the different games are ordered in the way they were played
data_long$game <- factor(data_long$game,levels=c("rps","fwg","shootout"))
ggplot(data_long,aes(x=round,y=probability,colour=strategy)) + geom_line() + facet_grid(game~stage)

```