### 2020-09-04 exp1_hmm_strategies.Rmd

---
title: "Hidden Markov strategy model"
author: "Maarten Speekenbrink"
date: "25/08/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(depmixS4)
```

```{r define-dummy-response}
setClass("dummyResponse", contains="response")

setGeneric("dummyResponse", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("dummyResponse"))

setMethod("dummyResponse", 
    signature(y="ANY"), 
    function(y,pstart=NULL,fixed=NULL, ...) {
      y <- matrix(y,length(y))
  		x <- matrix(1)
  		parameters <- list()
  		npar <- 0
      mod <- new("dummyResponse",parameters=parameters,fixed=logical(0),x=x,y=y,npar=npar)
      mod
	}
)

setMethod("show","dummyResponse",
    function(object) {
        cat("Dummy for fixed likelihood Model \n")
    }
)

setMethod("dens","dummyResponse",
  function(object,log=FALSE) {
   if(log) log(as.numeric(object@y)) else as.numeric(object@y)
  }
)

setMethod("getpars","dummyResponse",
    function(object,which="pars",...) {
        switch(which,
            "pars" = {
                pars <- numeric(0)
            },
            "fixed" = {
                pars <- logical(0)
            }
        )
        return(pars)
    }
)

setMethod("setpars","dummyResponse",
    function(object, values, which="pars", ...) {
        npar <- npar(object)
        if(length(values)!=npar) stop("length of 'values' must be",npar)
        # determine whether parameters or fixed constraints are being set
		nms <- ""
		switch(which,
		  "pars"= {
		      },
		  "fixed" = {
		    }
		  )
      names(object@parameters) <- nms
      return(object)
    }
)

setMethod("fit","dummyResponse",
    function(object,w) {
		  return(object)
	}
)

setMethod("predict","dummyResponse", 
    function(object) {
        ret <- object@y
        return(ret)
    }
)
```

```{r}
# we need to have "dat" available and lik_by_trial
QLS_lik <- lik_by_trial$QLS_lik
QLS_lik[which(QLS_lik == 0)] <- .001
QLS_lik[is.na(QLS_lik)] <- .33 # this is a hack; there shouldn't be any missing values
Bayes_NT_lik <- lik_by_trial$Bayes_NT_lik
Bayes_Tr_lik <- lik_by_trial$Bayes_Tr_lik

nsubject <- length(unique(dat$human_id)) # number of participants
ngame <- 3 # number of games
ntrial <- c(50,50,50) # numer of trials in each game

rModels <- list(
  list(
	  dummyResponse(QLS_lik)
	),
	list(
		dummyResponse(Bayes_NT_lik)
	),
  list(
    dummyResponse(Bayes_Tr_lik)
  )
)

trstart <- matrix(c(0.8,0.1,0.1,0.1,0.8,0.1,.1,.1,.8),ncol=3)
transition <- list()
transition[[1]] <- transInit(~1,nstates=3,data=data.frame(1),pstart=trstart[1,],family=multinomial("identity"))
transition[[2]] <- transInit(~1,nstates=3,data=data.frame(1),pstart=trstart[2,],family=multinomial("identity"))
transition[[3]] <- transInit(~1,nstates=3,data=data.frame(1),pstart=trstart[3,],family=multinomial("identity"))

instart <- c(1/3,1/3,1/3)
inMod <- transInit(~1,nstates=3,pstart=instart,family=multinomial("identity"),data=data.frame(rep(1,nsubject*ngame)))

mod <- makeDepmix(response=rModels,transition=transition,prior=inMod,ntimes=rep(ntrial,nsubject))

fmod <- fit(mod, emcontrol=em.control(random.start=FALSE))

# No switching. Force off diagonal initial elements of transtion matrix to 0
trstart <- matrix(c(1,0,0,0,1,0,0,0,1),ncol=3)
transition <- list()
transition[[1]] <- transInit(~1,nstates=3,data=data.frame(1),pstart=trstart[1,],family=multinomial("identity"))
transition[[2]] <- transInit(~1,nstates=3,data=data.frame(1),pstart=trstart[2,],family=multinomial("identity"))
transition[[3]] <- transInit(~1,nstates=3,data=data.frame(1),pstart=trstart[3,],family=multinomial("identity"))

mod_noswitch <- makeDepmix(response=rModels,transition=transition,prior=inMod,ntimes=rep(ntrial,nsubject))

fmod_noswitch <- fit(mod_noswitch, emcontrol=em.control(random.start=FALSE))

# p-value for comparison between a model with strategy switches and one without:

1-pchisq(-2*as.numeric(logLik(fmod_noswitch)) - (-2*as.numeric(logLik(fmod))),df=6)

```