### 2021-09-30 CBB/CBB.tex
CBB manuscript

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

% Pandoc citation processing

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\title{Transfer of Learned Opponent Models in Zero Sum Games \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }

\author{  Ismail Guennouni \and  Maarten Speekenbrink \and  }

\institute{
        Ismail Guennouni \at
     Department of Experimental Psychology, University College London, 26
 Bedford Way, London WC1H 0AP, United Kingdom \\
     \email{\href{mailto:i.guennouni.17@ucl.ac.uk}{\nolinkurl{i.guennouni.17@ucl.ac.uk}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Maarten Speekenbrink \at
     Department of Experimental Psychology, University College London, 26
 Bedford Way, London WC1H 0AP, United Kingdom \\
     \email{\href{mailto:m.speekenbrink@ucl.ac.uk}{\nolinkurl{m.speekenbrink@ucl.ac.uk}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Human learning transfer abilities take advantage of important cognitive
building blocks such as an abstract representation of concepts
underlying tasks and causal models of the environment. One way to build
abstract representations of the environment when the task involves
interactions with others is to build a model of the opponent that may
inform what actions they are likely to take next. In this study, we
explore opponent modelling and its transfer with the use of computer
agents with human-like limited degrees of iterated reasoning. In two
experiments, we find that participants deviate from Nash equilibrium
play and learn to adapt to their opponent's strategy to exploit it.
Moreover, we show that participants transfer their learning to new games
and that this transfer is moderated by the level of sophistication of
the opponent (second experiment), as well as the similarity between
games. Computational modelling shows that players start each game with a
model-based learning strategy that facilitates between-game transfer of
their opponent's strategy, but then switch to behaviour that is
consistent with a model-free learning strategy in the latter stages of
the interaction.
\\
\keywords{
        Learning Transfer \and
        Opponent Modelling \and
        Hidden Markov Models \and
        Bayesian Cognitive Hierarchy \and
    }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Being able to transfer previously acquired knowledge to a new domain is
one of the hallmarks of human intelligence. This ability relies on
important cognitive building blocks, such as an abstract representation
of concepts underlying tasks (Lake et al. 2017). One way to form these
representations when the task involves interactions with others, is to
build a model of the person we are interacting with that offers
predictions of the actions they are likely to take next. There is
evidence that people learn such models of their opponents when playing
repeated economic games (Stahl and Wilson 1995).

A model of the opponent can help increase performance in a particular
game, but learning more general characteristics of an opponent may also
help increase performance in other games. In this paper, we are
specifically interested in the latter: How do people build and use
models of their opponent to facilitate learning transfer? Repeated
games, in which players interact repeatedly with the same opponent and
have the ability to learn about their opponent's strategies and
preferences (Mertens 1990) are particularly useful to address this
question. The early literature on learning transfer in repeated games
has mostly focused on the proportion of people who play normatively
optimal (e.g.~Nash equilibrium play) or use salient (e.g.~risk
dominance) actions in later games, having had experience with a similar
game environment previously (Ho, Camerer, and Weigelt 1998; Camerer and
Knez 2000). As is well-known, a Nash equilibrium means that all players
in a game act such that no-one can unilaterally improve their
performance by deviating from their strategy. When playing against an
opponent with a Nash-optimal strategy, I can do no better than play
according to the Nash-equilibrium strategy as well. However, when faced
with a player who deviates from the Nash-optimal strategy, I may be able
to exploit this and also deviate from this strategy, increasing my
performance beyond what is expected at Nash equilibrium. Of course, this
comes with some risk, as my own deviation from Nash-optimal play may
leave me open to similar exploitation.

Studies that focused on whether people can learn to exploit deviations
from Nash-equilibrium play have mostly looked at the ability of players
to detect and exploit action-based learning rules (Spiliopoulos 2013;
Shachat and Swarthout 2004). These studies used computer opponents that
don't adapt play to their human opponent, mostly consisting of playing
each action with a fixed probability (a mixed strategy) or using a
pre-determined sequence. Findings showed that humans are capable of
adapting to non-Nash-equilibrium play and detect patterns in an
opponent's history of play. However, the use of mixed strategies may
have limited people's ability to form accurate opponent models.

Instead of thinking of their opponents as drawing random actions from a
(non-uniform) probability distribution, people may rather think of their
opponents as applying a form of iterated reasoning to determine their
next action. The type of reasoning we refer to takes the form of ``I
believe that you believe that I believe \ldots{}'' and has been shown to
underlie people's decisions in economic games (Camerer 2003;
Costa-Gomes, Crawford, and Broseta 2001). It also underlies successful
models in behavioural economics, such as Level-\(k\) and Cognitive
Hierarchy models (Camerer, Ho, and Chong 2004), which posit that people
assume their opponent applies a limited level of iterative reasoning,
taking actions that are best responses to those of their (modelled)
opponent.

In Level-\(k\) theory, a level-0 player uses a fixed strategy without
explicitly considering the strategy of their opponent. A level-1 player
assumes their opponent is a level-0 player, and chooses actions to best
respond to the strategy of their opponent, without considering what
their opponent might believe that they will play. A level-2 player, on
the other hand, takes their opponent's belief about their actions into
account, assuming they face a level-1 player, and choosing actions to
best respond to the actions of that player. Cognitive Hierarchy theory
is based on similar principles, but rather than assuming an opponent
always adopts a particular level-\(k\) strategy, they are assumed to
adopt each of the level-\(k\) strategies with a particular probability
(i.e., the opponent's strategy is a mixture over pure level-\(k\)
strategies).

Iterative reasoning strategies can explain non-equilibrium play in a
range of games, such as the \(p\)-beauty contest game (Nagel 1995),
dominance solvable games (Camerer, Ho, and Chong 2004), and various
standard normal form games (Costa-Gomes, Crawford, and Broseta 2001). In
the present study, we endow our computer opponents with a limited
ability for iterative reasoning and assess whether (1) human players
adapt their strategy to exploit this limited reasoning of their
opponent, and (2) whether they are able to generalize a learned opponent
model to other games. In two experiments, participants face the same
opponent (Experiment 1) or the same two opponents (Experiment 2) in
three consecutive games: the well-known Rock-Paper-Scissors game, the
structurally similar Fire-Water-Grass game, and a less similar Numbers
(Experiment 1) or Shootout (Experiment 2) game. To foreshadow our
results, we find evidence that participants transfer the learned
strategy of their opponent to other games, and that this transfer is
moderated by the sophistication of their opponent and the similarity
between the games. Moreover, using computational modelling, we find
evidence that participants switch from relying on an opponent model in
the early stages of the games, to a more habitual and cognitively less
demanding strategy in the later stages of the games.

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

In the first experiment, we aim to test learning transfer by making
participants face the same computer opponent with a limited level of
iterative reasoning in three sequential games that vary in similarity.
If participants are able to learn the limitations of their opponent's
iterative reasoning and generalize this to new games, their performance
in (the early stages of) later games should be higher than expected if
they were to completely learn a new strategy in each game.

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{participants-and-design}{%
\subsubsection{Participants and Design}\label{participants-and-design}}

A total of 52 (28 female, 24 male) participants were recruited on the
Prolific Academic platform. The mean age of participants was 31.2 years.
Participants were paid a fixed fee of £2.5 plus a bonus dependent on
their performance (£1.06 on average). The experiment had a 2 (computer
opponent: level 1 or level 2) by 3 (games: rock-paper-scissors,
fire-water-grass, numbers) design, with repeated measures on the second
factor. Participants were randomly assigned to one of the two levels of
the first factor.

\hypertarget{tasks}{%
\subsubsection{Tasks}\label{tasks}}

Participants played the three games against their computer opponent.
These games were Rock-Paper-Scissors, Fire-Water-Grass, and the Numbers
game. A typical Rock-Paper-Scissors game (hereafter RPS) is a 3x3
zero-sum game, with a cyclical hierarchy between the two player's
actions: rock blunts scissors, paper wraps rock, and scissors cut paper.
If one player chooses an action which dominates their opponent's action,
the player wins (receives a reward of 1) and the other player loses
(receives a reward of -1). Otherwise it is a draw and both players
receive a reward of 0. RPS has a unique mixed-strategy Nash equilibrium,
which consists of each player in each round randomly selecting from the
three options with uniform probability.

The Fire-Water-Grass (FWG) game is identical to RPS in all but action
labels: Fire burns grass, water extinguishes fire, and grass absorbs
water. We use this game as we are interested in whether learning is
transferred in a fundamentally similar game where the only difference is
in the label of the possible actions. This should make it relatively
easy to generalize knowledge of the opponent's strategy, provided this
knowledge is on a sufficiently abstract level, such as knowing the
opponent is a level-1 or level-2 player. Crucially, learning simple
contingencies such as ``If I played rock on the previous round, playing
scissors next will likely result in a win'', is not generalizable to
this similar game, as these contingencies are tied to the labels of the
actions.

The Numbers game is a generalization of RPS. In the variant we use, 2
participants concurrently pick a number between 1 and 5. To win in this
game, a participant needs to pick a number exactly 1 higher than the
number chosen by their opponent. For example, if a participant thinks
their opponent will pick 3, they ought to choose 4 to win the round. To
make the strategies cyclical as in RPS, the game stipulates that the
lowest number (1) beats the highest number (5), so if the participant
thinks the opponent will play 5, then the winning choice is to pick 1.
This game has a structure similar to RPS in which every action is
dominated by exactly one other action. All other possible combinations
of choices are considered ties. Similar to RPS and FWG, the
mixed-strategy Nash equilibrium is to randomly play each action with
equal probability.

The computer opponent was programmed to use either a level-1 or level-2
strategy in all the games. A level-1 player is defined as a player who
best responds to a level-0 player. A level-0 player plays in a
non-strategic way and does not consider their opponent's actions. Here,
we assume a level-0 player simply repeats their previous action. There
are other ways to define a level-0 player. For instance, as repeating
their action if it resulted in a win and choosing randomly from the
remaining actions otherwise, or choosing randomly from all actions. As a
best response to a uniformly random action is itself a random action,
defining a level-0 player in such a way would make a level-1 opponent's
strategy much harder to discern. Because we are mainly interested in
generalization of knowledge of an opponent's strategy to other games,
which requires good knowledge of this strategy, we opted for this more
deterministic formulation of a level-0 player. The level-2 computer
opponent assumes in turn that the participant is a level-1 opponent,
playing according to the strategy just described. To make the computer
opponent's strategy not too obvious, we introduced some randomness in
actions of computer opponent, making them play randomly in 10\% of all
trials. Note that at all levels, the strategies are contingent on the
actions taken in the previous round. The choice of this type of strategy
is consistent with evidence that humans strategically use information
from last round play of their opponents in zero-sum games (Batzilis et
al. 2016; Wang, Xu, and Zhou 2014). Table \ref{tab:table-actions} shows
an example of the computer opponent's actions in response to the
previous round play.

\begin{table}

\caption{\label{tab:table-actions}Example of how a level-1 and level-2 computer agent plays in response to actions taken in the previous round}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llll}
\toprule
Human action t-1 & Computer action t-1 & Computer level-1 action t & Computer level-2 action t\\
\midrule
Paper & Rock & Scissors & Scissors\\
Scissors & Scissors & Rock & Paper\\
Rock & Paper & Paper & Rock\\
... & ... & ... & ...\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{procedure}{%
\subsubsection{Procedure}\label{procedure}}

Participants were informed they would play three different games against
the same computer opponent. Participants were told that the opponent
cannot cheat and will choose its actions simultaneously with them,
without prior knowledge of the participant's choice. After providing
informed consent and reading the instructions, participants answered a
number of comprehension questions. They then played the three games
against their opponent in the order RPS, FGW, and Numbers. An example of
the interface for the RPS game is provided in Figure
\ref{fig:feedback-rps-exp2}. On each round, the human player chooses an
action, and after a random delay (between 0.5 and 3 seconds) is shown
the action chosen by the computer opponent and the outcome of that
round. A total of 50 rounds of each game was played with the player's
score displayed at the end of each game. The score was calculated as the
number of wins minus the number of losses. Ties did not affect the
score. In order to incentivise the participants to maximise the number
of wins against their opponent, players were paid a bonus at the end of
the experiment proportional to their final score (each point was worth
£0.02). After playing all the games, participants were asked questions
about their beliefs about the computer opponent, related to whether they
thought they learned their opponent's strategy, and how difficult they
found playing against their opponent. They were then debriefed and
thanked for their participation.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{../Report/images/feedback_rps}

}

\caption{Screenshot of the Rock-Paper-Scissors game in Experiment 2. Shown here is the feedback stage, after both the human (left) and computer (right) players have chosen their action. The interface was similar in Experiment 1, but excluded the history of game play in the center panel.}\label{fig:feedback-rps-exp2}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp1-avg-scores-game-1}

}

\caption{Performance per game and block across conditions in Experiment 1. Points are scores of individual participants and boxes reflect the 95\% confidence intervals of the mean (center line equals the mean).}\label{fig:exp1-avg-scores-game}
\end{figure}

\hypertarget{results}{%
\subsection{Results}\label{results}}

First, we focus on participants scores in each game (see Figure
\ref{fig:exp1-avg-scores-game}). On average, participants obtained the
lowest score in the RPS game (\(M = 0.289\), \(SD = 0.348\)), followed
by Numbers (\(M = 0.31\), \(SD = 0.347\)). Participants' performance was
highest in the FWG game (\(M = 0.454\), \(SD = 0.354\)). Scores in each
game were significantly different from 0, the expected score of
uniformly random play (RPS: \(t(51) = 7.26\), \(p < .001\); FWG:
\(t(51) = 10.04\) , \(p < .001\); Numbers: \(t(51) = 7.17\),
\(p < .001\)). As uniformly random play is the Nash equilibrium, this
indicates successful deviation from a Nash-optimal strategy.

To assess learning within and between games, we estimate a linear
mixed-effects model with score as the dependent variable, and Condition
(Level 1 or Level 2), Game (RPS, FWG, Numbers) and Block (first-half ,
second-half) as independent variables, including all main effects and
interactions. The model included random intercepts for participants. The
results show a main effect of Game (\(F(2,250) = 12.90\),
\(\eta^{2} = 0.09\), \(p < .001\)). Post-hoc pairwise comparisons show
that performance in the FWG game was significantly higher than in the
RPS game (\(t(100) =3.78\), \(p < .001\)) and the Numbers game
(\(t(100) = 3.32\) , \(p = .002\)). The score in RPS was not
significantly different from the score in Numbers (\(t(100) = 0.45\) ,
\(p = .65\)). The main effect of Block (\(F(2,250) = 15.16\) ,
\(\eta^{2} = 0.06\), \(p < .001\)) shows that the score in the first
half of each game (\(M = 0.29\)) was significantly lower than in the
second half (\(M = 0.40\)), which indicates within-game learning. The
main effect of Condition (\(F(1,50) = 5.44\), \(\eta^{2} = 0.10\),
\(p = .024\)) indicates that scores were higher against the level-1
player (\(M = 0.43\)) than against the level-2 player (\(M = 0.27\)).
Thus, it appears that is was harder for participants to exploit the
strategy of the more sophisticated level-2 opponent than the
comparatively less sophisticated level-1 opponent. There was also an
interaction between Game and Condition (\(F(2,250) = 4.52\) ,
\(\eta^{2} = 0.03\), \(p = .01\)). Post-hoc pairwise comparisons
(\(p\)-values adjusted using Holm method for multiple comparisons) show
that, for participants facing level-1 opponents, scores on the FWG game
were significantly higher than those for both RPS (\(t(250) = 3.85\) ,
\(p = .002\)), and Numbers (\(t(250) = 3.79\) , \(p = .002\)). For
participants facing level-2 opponents however, there was no significant
difference between scores on the three games.

\hypertarget{learning-transfer}{%
\subsection{Learning transfer}\label{learning-transfer}}

As a measure for learning transfer, we focus on participants' scores in
the initial 5 rounds after the first round (rounds 2-6) of each game
(see Figure \ref{fig:exp1-early-score-by-opp}). We exclude the very
first round as the computer opponent plays randomly here and there is no
opportunity yet for the human player to exploit their opponent's
strategy. Players with no knowledge of their opponent's strategy are
expected to perform at chance level in these early rounds. Positive
scores in rounds 2-6 reflect generalization of prior experience. The FWG
early score score is significantly higher than 0 (\(t(148.85) = 4.584\),
\(p < .001\)). This is also the case for the Numbers game
(\(t(148.85) = 3.00\), \(p = .009\)). We did not expect positive scores
for the RPS game, as it was the first game played and there was no
opportunity for learning about the opponent's strategy. Scores in this
game were indeed not significantly different from 0
(\(t(148.85) = 1.04\) , \(p = .89\)).

\begin{figure}

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp1-early-score-by-opp-1}

}

\caption{\label{ref:figure4-caption}Performance in early rounds (2-6) per game and block across conditions in Experiment 1. Points are scores of individual participants and boxes reflect the 95\% confidence intervals of the mean (center line equals the mean).}\label{fig:exp1-early-score-by-opp}
\end{figure}

Next, we explore whether learning transfer is moderated by the type of
opponent and game similarity. We expected better transfer between more
similar games (i.e.~better transfer from RPS to FWG than from RPS/FWG to
Numbers), and worse transfer for the more sophisticated level-2 agent.
Figure \ref{fig:exp1-early-score-by-opp} indicates that the pattern over
the games is indeed dissimilar between level-1 and level-2 opponents. We
estimates a linear mixed-effects model with the early round score as the
dependent variable, and Condition (Level 1 or Level 2) and Game (RPS,
FWG, Numbers), as well as their interaction, as independent variables.
The model included random intercepts for participants. The results
indicated a main effect of Game (\(F(2,100) = 3.35\), \(p = .04\)), but
no effect for condition. Further investigation of the effects of game on
early score using planned contrasts revealed that early scores in the
FWG games are significantly higher than in the RPS (
\(t(99.99) = -2.32\), \(p = 0.02\)).

We next ran statistical tests on early round scores by game and
opponent, against the null hypothesis of 0 (no transfer). For level-1
facing players, there is evidence of learning transfer from RPS to both
FWG (\(t(150) = 3.96\), \(p < .001\)) and Numbers (\(t(150) = 3.74\),
\(p < .001\)). For level-2 facing players, there is evidence for
transfer from RPS to the similar game FWG, albeit scores are lower than
for level-1 player (\(t(150) = 2.48\), \(p = .01\)) but not to the
dissimilar game of Numbers.

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

The results of Experiment 1 indicate that participants were able to
learn successful strategies which exploited the deviation from
Nash-optimal play of their opponents. Moreover, they were able to
transfer knowledge about their opponent to other games, but this was
moderated by the type of game. There was evidence of transfer from the
RPS game to the similar FWG game for both opponents. However, transfer
to the dissimilar Numbers game was only evident for participants facing
a level-1 opponent.

\hypertarget{experiment-2}{%
\section{Experiment 2}\label{experiment-2}}

In Experiment 2, we aimed to obtain a stronger test of learning
transfer. Instead of facing a single level-1 or level-2 opponent
throughout all games, participants now faced both types of opponent. To
perform well against both opponents, participants would need to learn
distinct strategies against these opponents. To reduce effects of
increased memory load due to facing distinct opponents, we provided
participants access to the history of play against an opponent within
each game (see Figure \ref{fig:feedback-rps-exp2}). Finally, we changed
the third game to a penalty Shootout game, with participants aiming to
score a goal and opponents playing the role of goal keepers. Whilst this
game has the same number of actions as the first two (aim left, center,
or right), it is strategically dissimilar. Unlike the Numbers game in
Experiment 1, the Shootout game does not have a cyclical hierarchy
between actions, making it harder to win through a heuristic based on
this cyclicity.

\hypertarget{methods-1}{%
\subsection{Methods}\label{methods-1}}

\hypertarget{participants-design}{%
\subsubsection{Participants \& Design}\label{participants-design}}

A total of 48 participants (21 females, 28 males, 1 unknown) were
recruited via the Prolific platform, none of which took part in
Experiment 1. The average age was 30.2 years, and the mean duration to
complete the task was 39 minutes. Participants received a fixed fee of
£2.5 for completing the experiment and a performance dependent bonus
(£1.32 on average).

\hypertarget{tasks-1}{%
\subsubsection{Tasks}\label{tasks-1}}

Participants played three games: Rock-Paper-Scissors (RPS),
Fire-Water-Grass (FWG), and the penalty Shootout game. The first two
games were identical to the ones used in Experiment 1. In the Shootout
game, participants took the role of the a football (soccer) player in a
penalty situation, with the computer opponent taking the role of the
goalkeeper. Players had the choice between three actions: shooting the
football to the left, right, or centre of the goal. Similarly, the
goalkeeper chooses between defending the left, right, or centre of the
goal. If participants shoot in a different direction than where the
goalkeeper defends, they win the round and the goalkeeper loses.
Otherwise, the goalkeeper catches the ball and the player loses the
round. There is no possibility of ties in this game. Figure
\ref{fig:screenshot-shootout} shows a snapshot of play in the Shootout
game. What makes this game different to the other games is that there
are now two ways to beat the opponent: if the shooter thinks their
opponent is going to choose to defend ``right'' in the next round, they
can win by either choosing to shoot ``left'' or ``center''. A level-1
shooter who thinks that their goalkeeper opponent will repeat their last
action has thus two possible best responses. A level-1 goalkeeper,
however, has only a single best response (defending where their opponent
aimed in the last round). A level-2 goalkeeper, who believes their
opponent is a level-1 shooter, will have two best responses however. We
programmed the level-2 computer player to choose randomly between these
two best responses.

As in Experiment 1, all the games have a unique mixed-strategy Nash
equilibrium consisting of uniformly random actions. If participants
follow this strategy, or simply don't engage in learning how the
opponent plays, they would score 0 on average against both level-1 and
level-2 players. Evidence of sustained wins would indicate that
participants have learned to exploit patterns in their opponents' play.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{../Report/images/shootout}

}

\caption{Screenshot of the shootout game}\label{fig:screenshot-shootout}
\end{figure}

\hypertarget{procedure-1}{%
\subsubsection{Procedure}\label{procedure-1}}

Participants played 3 games sequentially against both level-1 and
level-2 computer opponents. As in Experiment 1, the computer opponents
retained the same strategy throughout the 3 games. Participants faced
each opponent twice in each game. Each game was divided into 4 stages,
numbered 1 to 4, consisting of 20, 20, 10, and 10 rounds respectively
for a total of 60 rounds per game. Participants started by facing one of
the opponents in stage one, then the other in stage two. This was
repeated in the same order in stages 3 and 4. Which opponent they faced
first was counterbalanced. All participants engaged in the three games
(RPS, FWG and Shootout) in this exact order, and were aware that their
opponent could not cheat and chose their action simultaneously with the
player, without knowing their choices beforehand. In order to encourage
participants to think about their next choice, a countdown timer of 3
seconds was introduced at the beginning of each round. During those 3
seconds, participants could not choose an action and had to wait for the
timer to run out. A random delay between 0.5 and 3 seconds was again
introduced before the choice of the computer agent was revealed, as a
way of simulating a real human opponent's decision time. After each
round, participants were given detailed feedback about their opponent's
action and whether they won or lost the round. Further information about
the history of play in previous rounds was also provided and
participants could scroll down to recall the full history of each
interaction against an opponent in a particular stage of a game. The
number of wins, losses and ties in each game were displayed at the top
of the screen, and this scoreboard was reinitialised to zero at the
onset of a new stage game.

\hypertarget{results-1}{%
\subsection{Results}\label{results-1}}

Participants' scores are depicted in Figure \ref{fig:exp2-score-by-opp}.
The RPS game had the lowest average score per round (\(M = 0.194\), SD =
0.345) followed by FWG (\(M = 0.27\), SD = 0.394) and finally the
Shootout game (\(M = 0.289\), SD = 0.326).\footnote{A higher score in
  the Shootout game is expected as there are 2 out of three possible
  winning actions, compared to one out of three in RPS and FWG. Indeed,
  a player not aiming to uncover the opponent's strategy and thus
  choosing to play randomly should be expected to have on average score
  per round of 0 in both RPS and FWG, and 0.33 in the Shootout game. To
  make the scores more comparable, and because we are interested in
  player's performance that is not due to chance, we will adjust all
  scores in the Shootout game by subtracting the average score per round
  of a random strategy (0.33).} Using one-sample t-tests on adjusted
scores, we reject the null hypothesis of random play in all three games
(RPS: \(t(49) = 6.26\), \(p < .001\); FWG: \(t(49) = 7.25\),
\(p < .001\); Shootout: \(t(49) = 13.61\), \(p < .001\)).

\begin{figure}

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp2-score-by-opp-1}

}

\caption{\label{fig:exp2-score-by-opp}Performance per game and interaction across opponents in Experiment 2. Points are scores of individual participants and boxes reflect the 95\% confidence intervals of the mean (center line equals the mean).}\label{fig:exp2-score-by-opp}
\end{figure}

To explore whether learning occurred within and between games, we
estimated a linear mixed-effects model, with score as the dependent
variable and Condition (level-1 first, level-2 first), Opponent type
(level-1 or level-2), Game (RPS, FWG, Shootout) and Encounter (first or
second) as the independent variables, including all main effects and
interactions. The model included random intercepts for participants. We
found a main effect of Game (\(F(2,528) = 15.32\), \(\eta^{2} = 0.04\),
\(p < .001\)). Post-hoc pairwise comparisons between games (\(p\)-values
adjusted using Holm method for multiple comparisons) indicate
performance in the games increased steadily throughout the experiment,
with FWG performance significantly higher than RPS (\(t(96) =2.53\),
\(p = .025\)), and performance in the Shootout game significantly higher
than in FWG (\(t(96) = 2.32\), \(p = .025\)). Whilst the model shows no
main effect of Condition, Opponent type, or Encounter, there was a
significant interaction between Game and Opponent type
(\(F(2, 528) = 5.78\) ,\(\eta^{2} = 0.02\), \(p = .003\)). Follow-up
analysis differentiating performance by type of opponent faced shows
that when facing level-1 agents, scores increased steadily after each
game, with FWG score significantly higher than RPS (\(t(191) = 2.70\),
\(p = .03\)) and Shootout scores in turn significantly higher than FWG
(\(t(191) = 3.05\), \(p = .01\)). There was no significant difference
between scores on games when facing level-2 agents however.

\hypertarget{learning-transfer-1}{%
\subsection{Learning transfer}\label{learning-transfer-1}}

To measure transfer between games, we again focus on participants'
scores in the initial 5 rounds after the first round (rounds 2-6) of
each game and only during the first time they interact with an opponent
in a game (see Figure \ref{fig:exp2-early-score-by-opp}). We estimated a
linear mixed-effects model with the early round scores as the dependent
variable, and Condition (level-1 first, level-2 first), Opponent type
(level-1 or level-2) and Game (RPS, FWG, Shootout) as independent
variables, including all main effects and interactions. The model also
included random intercepts for participants. We found a main effect of
Game (\(F(2,240) = 6.57\), \(p = .002\)). The effect of game reflects
early scores on FWG being significantly higher than on RPS
(\(t(240) = 3.01\), \(p = .005\)) as well as scores on shootout being
significantly higher than RPS (\(t(240) = 3.25\), \(p = .003\)). We also
found an interaction effect between Game and Opponent type
(\(F(2,240) = 3.34\), \(p = .037\)). Post-hoc pairwise comparisons
(p-values adjusted using Holm method for multiple comparisons) show that
the difference between early scores in RPS and shootout was mostly
driven by participants facing level-1 players ((\(t(240) = 4.12\),
\(p < .001\))). No signifcant differences were found between
participants facing different opponents in the same game, or between
games when facing the same opponent.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp2-early-score-by-opp-1}

}

\caption{\label{ref:figure4-caption}Performance in early rounds (2-6) per game and opponent in Experiment 2. Points are scores of individual participants and boxes reflect the 95\% confidence intervals of the mean (center line equals the mean).}\label{fig:exp2-early-score-by-opp}
\end{figure}

We then ran statistical tests on early round scores by game and opponent
against the null hypothesis of 0 (no transfer). For level-1 facing
players, there is evidence of learning transfer from RPS to both FWG
(\(t(271) = 4.99\), \(p < .001\)) and Shootout (\(t(271) = 6.66\),
\(p < .001\)). For level-2 facing players, there is evidence for
transfer from RPS to the similar game FWG, albeit scores are lower than
for level-1 player (\(t(271) = 4.41\), \(p < .001\)). Unlike in
Experiment 1, there is also evidence of transfer to the dissimilar
Shootout game (\(t(271) = 3.22\), \(p = .004\)). The early round scores
in the similar FWG games for level-1 and level-2 facing players are not
significantly different from each other. However, the score of the
players facing the level-1 opponent in Shootout is higher than that of
players facing level-2 opponents (\(t(144) = 2.45\), \(p = 0.01\)).

\hypertarget{discussion-1}{%
\subsection{Discussion}\label{discussion-1}}

The results of Experiment 2 confirm earlier findings from Experiment 1
on learning transfer in a situation where participants need to learn
about two distinct opponents. As in Experiment 1, participants adapted
their strategies to both level-1 and level-2 opponents to exploit
deviations from Nash-optimal play. Knowledge about both types of
opponents were transferred to both the similar and dissimilar game. This
transfer was moderated by the degree of similarity between games, as
well as opponent type. The interaction effect showed that transfer to
the dissimilar game was easier when facing the less sophisticated agent.

\hypertarget{computational-modelling}{%
\section{Computational modelling}\label{computational-modelling}}

To gain more insight into participants' strategies against their
computer opponents, we constructed and tested several computational
models of strategy learning. The baseline model assumes play is random,
and each potential action is chosen with equal probability. Note that
this corresponds to the Nash equilibrium strategy. The other models
adapted their play to the opponent, either by reinforcing successful
actions in each game (reinforcement learning), or by determining the
type of opponent through Bayesian learning (Bayesian Cognitive Hierarchy
models). We also include the (self-tuning) Expected Weighted Attraction
(EWA) model, which is a popular model in behavioural economics.

In the following, we will describe the models in more detail, and
provide some intuition into how they implement learning about the game
and/or the opponent. Throughout, we use the following notation: In each
game
\(g \in \{\text{RPS},\text{FWG}, \text{Numbers}, \text{Shootout} \}\),
on each trial \(t\), the participant chooses an action
\(a_t \in \mathcal{A}_g\), and the opponent chooses action
\(o_t \in \mathcal{A}_g\), where \(\mathcal{A}_g\) is the set of allowed
actions in game \(g\), e.g. \(\mathcal{A}_\text{RPS} = \{R,P,S\}\). The
participant then receives reward \(r_t \in \{1,0,-1\}\), and the
opponent receives \(-r_t\). We use the state variable
\(s_t = \{a_{t-1},o_{t-1}\}\) to denote the actions taken in the
previous round \(t-1\) by the participant and opponent. The initial
state is empty, and we assume that the action at a first encounter of an
opponent in a game is chosen at random.

\hypertarget{reinforcement-learning-rl-model}{%
\subsection{Reinforcement learning (RL)
model}\label{reinforcement-learning-rl-model}}

We first consider a model-free reinforcement learning algorithm, where
actions that have led to positive rewards are reinforced, and the
likelihood of actions that led to a negative reward is lowered. Since
the computer players in this experiment based their play on the actions
in the previous round, a suitable RL model for this situation is one
which learns the value of actions contingent on plays in the previous
round, i.e.~by defining the state \(s_{t}\) as above. The resulting RL
model learns a \(Q\)-value (Watkins and Dayan 1992) for each
state-action pair:
\[Q_{t+1}(s_{t},a_{t}) = Q_{t}(s_{t},a_{t}) + \alpha \left( r_{t}  - Q(s_{t},a_{t}) \right) ,\]
where \(Q(s_{t},a_{t})\) is the value of taking action \(a\) when in
state \(s\) at time \(t\), and \(\alpha \in [0,1]\) the learning rate.
For instance, \(Q_t(\{R,S\},P)\) denotes the value of taking action
``Paper'' this round if the player's last action was ``Rock'' and the
opponent played ``Scissors''. Actions are taken according to a softmax
rule:
\[P_{t}(a|s_t) = \frac{\exp \{ \lambda Q_{t}(a,s_t) \}}{\sum_{a' \in \mathcal{A}_g} \exp \{\lambda  Q_{t}(a',s_t) \}}, \]
where the inverse temperature parameter \(\lambda\) determines the
consistency of the strategy (the higher \(\lambda\), the more often the
action with the highest \(Q\)-value is chosen). While this RL model
allows the players to compute the values of actions conditional on past
play, crucially, it will not be able to transfer learning between games,
as each game has a different state space \(\mathcal{S}_g\) and action
space \(\mathcal{A}_g\), and there is no simple way to map states and
actions across games. The RL model has two free parameters: the learning
rate (\(\alpha\)) and the inverse temperature (\(\lambda\)).

\hypertarget{experience-weighted-attraction-ewa-model}{%
\subsection{Experience-weighted attraction (EWA)
model}\label{experience-weighted-attraction-ewa-model}}

The self-tuning Experience Weighted Attraction (EWA) model (Ho, Camerer,
and Chong 2004) combines two seemingly different approaches, namely
reinforcement learning and belief learning. Belief learning models are
based on the assumption that players keep track of their opponent's
frequency of past actions, and best respond to that. By contrast,
reinforcement learning does not explicitly take into account beliefs
about other players, but simply increases the probability of repeating a
more rewarding action. The self-tuning EWA model has been shown to
perform better than either RL or belief learning alone in various
repeated games and has the advantage of having only one free parameter,
the inverse temperature of the softmax choice function. The EWA model is
based on updating ``Attractions'' for each action over time given a
particular state. The attraction of action \(a\) at time \(t\) given
state \(s\) is denoted as \(A_{t}(a, s)\) and updated as
\[ A_{t+1}(a,s) =  \frac{\phi(t) \ N(t) \ A_{t}(a,s) + [ \delta_{a}(t) + (1-\delta_{a}(t)) \ I(a_t = a )] \ R(a,o_t) } {\phi(t)N(t) + 1} \]
where \(I(x)\) is an indicator function which takes the value 1 if its
argument is true and 0 otherwise, and \(R(a,o_t)\) is the reward that
would be obtained from playing action \(a\) against opponent action
\(o_t\). \(R(a,o_t)\) equals the actual obtained reward when
\(a = a_t\), and otherwise is a counterfactual reward that would have
been obtained if a different action were taken. Unlike reinforcement
learning, this uses knowledge of the rules of the game to allow
reinforcing actions that were not actually taken by the rewards they
would have provided. The parameter \(\delta\) reflects the weight given
to such counterfactual rewards. Setting \(\delta = 0\) leads to
reinforcement only of actions taken, while positive values of \(\delta\)
makes the update rule take into account foregone payoffs, which is
similar to weighted fictitious play (Cheung and Friedman 1994). \(N(t)\)
represents an experience weight and can be interpreted as the number of
``observation-equivalents'' of past experience. We initialise it to 1 so
initial attractions and reinforcement from payoffs are weighted equally.

In the earlier version of the EWA model (Camerer, Ho, and Others 1997),
\(\phi\) and \(\delta\) were time-invariant free parameters. In the
self-tuning EWA model (Ho, Camerer, and Chong 2004), the values of
\(\delta\) and \(\phi\) are learnt from experience. Over time,
\(\delta\) is updated as
\[\delta_{a}(t) = \begin{cases} 1 & \text{if }  R(a,o_{t}) \geq r_{t}  \\
0 & \text{otherwise} \end{cases}\] The \(\phi_{t}\) parameter can be
interpreted as a discount of prior experience, modelling either limited
agent memory or changes in the game conditions. At the core, \(\phi(t)\)
depends on a surprise index \(S_{p}(t)\):
\[\phi(t) = 1 - \frac{1}{2}S_{p}(t) ,\] where \(S_{p}(t)\) quantifies
how the opponent deviates from past play. It is calculated in turn
through the cumulative history of play, across opponent possible actions
k, as \(h^{k}(t)\):
\[h^{k}(t)= \frac{ \sum_{\tau = 1}^t  I( o_{\tau} = o^k )} {t}, \] as
well as an indicator vector of the most recent opponent play:
\[r^k(t) = I(o^k=o_{t}), \] Where \(I\) is the indicator function as
defined above. To get the surprise index, we simply sum all the squared
deviations between the cumulative history vector \(h^{k}(t)\) and the
immediate history \(r^k(t)\):

\[S_{p}(t) = \sum_{k=1}^{|\mathcal{A}_g|} (h^{k}(t) - r^k(t))^2 \]

For more details on the self-tuning EWA model, we refer the reader to
(Ho, Camerer, and Chong 2004). As in the RL model above, actions are
chosen based on a softmax decision rule:

\[P_t(a,s) = \frac{\exp \{\lambda  A_{t}(a,s) \} }{\sum_{a' \in \mathcal{A}_t} \exp \{ \lambda A_{t}(a',s) \} } \]
The self-tuning EWA has one free parameter: the inverse temperature of
the softmax decision rule (\(\lambda\)).

\hypertarget{bayesian-cognitive-hierarchy-bch-model}{%
\subsection{Bayesian Cognitive Hierarchy (BCH)
model}\label{bayesian-cognitive-hierarchy-bch-model}}

In what we call the Bayesian Cognitive Hierarchy (BCH) model, the
participant attempts to learn the type of opponent they are facing
through Bayesian learning. For present purposes, we assume participants
consider the opponent could be either a level 0, level 1, or level 2
player, and start with a prior belief that each of these types is
equally likely. They then use observations of the opponent's actions to
infer a posterior probability of each type:
\[P(\text{level}=k | \mathcal{D}_{t})  \propto  P(\mathcal{D}_{t}|\text{level}=k ) \times P(\text{level}=k)\]
where \(\mathcal{D}_{t} = \{s_1,\ldots,s_t\}\) is the data available at
time \(t\). The likelihood is defined as
\[P(\mathcal{D}_{t}|\text{level}=k) = \prod_{j=1}^t \left( \theta \frac{1}{|\mathcal{A}_g|} + (1-\theta) f_k(o_j|s_{j})\right)\]
where \(f_k(o_t|s_{t}) = 1\) if \(o_t\) is the action taken by a level
\(k\) player when the previous round play was
\(s_t = (a_{t-1}, o_{t-1})\), and 0 otherwise. Note that the likelihood
assumes (correctly) that there is a probability \(\theta \in [0,1]\)
that the opponent takes a random action. The posterior at time \(t-1\)
forms the prior at time \(t\). We assume a participant chooses an action
by using the softmax function over the best response to predicted
actions:
\[\begin{aligned} B_t(a) &= \sum_{k = 0}^2 \sum_{o \in \mathcal{A}_g} b(a,o) f_k(o|s_{t})  P(\text{level}=k|\mathcal{D}_{t-1})\\
P_t(a) &= \frac{\exp\{\lambda B_t(a) \}}{\sum_{a' \in \mathcal{A}_g} \exp \{ \lambda B_t(a')\}} \end{aligned}\]
where \(b(a,o) = 1\) if action \(a\) is a best response to opponent's
action \(o\) (i.e.~it leads to a win).

Unlike the models above, the BCH model allows for between-game transfer,
as knowledge of the level of the opponent can be used to generate
predictions in games that have not been played before. This
generalization is done simply by using the posterior
\(P(\text{level} = k|\mathcal{D}_T)\) from the previous game as the
prior distribution in the next game. However, the participant might also
assume that the level of reasoning of their opponent does not generalize
over games. This would mean starting with a ``fresh'' prior
\(P(\text{level} = k)\) at the start of each game. We hence distinguish
between two versions of the BCH model. In the No-Between-Transfer
(BCH\_NBT) variant, participants assume a uniform probability of the
different levels at the start of each game (and hence do not transfer
knowledge of their opponent between games). In the Between-Transfer
model (BCH\_BT), participants use the posterior probability over the
levels of their opponent as the prior at the start of a new game
(i.e.~complete transfer of the knowledge of their opponent). Both
versions of the BCH model have two free parameters: the assumed
probability that the opponent chooses a random action (\(\theta\)), and
the temperature parameter of the softmax function (\(\lambda\)).

\hypertarget{estimation-and-model-comparison}{%
\subsection{Estimation and model
comparison}\label{estimation-and-model-comparison}}

For both experiments, we fitted all models to individual participant
data by maximum likelihood estimation using the DEoptim R package
(Mullen et al. 2011). We use the Bayesian Information Criterion (BIC) to
determine the best fitting model for each participant. For Experiment 1,
we fitted a total of 5 models: a baseline model assuming random play
(Nash), the Bayesian Cognitive Hierarchy model allowing transfer between
games (BCH\_BT) and without transfer between games (BCH\_NT), as well as
a reinforcement learning (RL), and finally a self-tuning EWA model with
the same state space (EWA). In Experiment 2, because participants were
interacting with each opponent twice within each game, we need to
distinguish between two types of opponent model transfer. We can have
transfer \emph{within} games between the first and second interaction
with the opponent. In addition, we can also have transfer \emph{between}
games, as in e.g.~transferring a learned opponent model from RPS to FWG.
Therefore, we fitted a total of three versions of the Bayesian Cognitive
Hierarchy model: BCH\_BT allows for both within and between game
transfer (between game transfer without within game transfer is
implausible); BCH\_WT allows only for within-game transfer, but not
between game transfer; BCH\_NT allows for no transfer within or between
games. As the RL model can't account for between game transfer due to a
change in state and action space, we can only have models that allowing
for within game transfer (RL\_WT) or with no transfer within games
(RL\_NT). Likewise, we fit both a self tuning EWA model with (EWA\_WT)
and without (EWA\_NT) within-game transfer. Including the baseline model
of random play (Nash), we therefore fitted a total of 8 models for
Experiment 2.

Figure \ref{fig:exp1-comp-models} shows the results for Experiment 1. We
can see that the RL model clearly described most participants' behaviour
best, followed by the random (Nash) model. Only a few participants were
best described by one of the BCH models or the EWA model.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp1-comp-models-1}

}

\caption{Experiment 1 - Histogram of best fitting computational models by condition}\label{fig:exp1-comp-models}
\end{figure}

In Experiment 2, we can see from Figure \ref{fig:exp2-comp-models} that
the RL model was again more successful than the Bayesian Cognitive
Hierarchy models (with or without transfer) and the EWA model in fitting
participants' action choices.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp2-comp-models-1}

}

\caption{Experiment 2 - Histogram of best fitting computational models by condition.}\label{fig:exp2-comp-models}
\end{figure}

\hypertarget{using-hidden-markov-model-to-explore-strategy-switching}{%
\subsection{Using Hidden Markov Model to explore strategy
switching}\label{using-hidden-markov-model-to-explore-strategy-switching}}

The computational modelling indicates that most players are best fit by
a reinforcement learning which learns good actions conditional upon the
last round play. This is at odds with the behavioural findings, where we
found evidence of transfer in early rounds of each game. If indeed most
participants adopt a reinforcement learning strategy, they should not be
able to transfer their learning to the early rounds of a new game. In
order to investigate this discrepancy further, we plot the likelihood by
trial for each game and three strategies: reinforcement learning (RL),
Bayesian Cognitive Hierarchy with between-game transfer (BCH\_BT), and
the random (Nash) strategy. Figure \ref{fig:exp1-lik-by-tr} shows that
in the initial rounds of the the later games in Experiment 1, the
likelihood for the BCH models is higher than that of the other models.
However, over time, the likelihood of the RL model increases and exceeds
that of the BCH model.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp1-lik-by-tr-1}

}

\caption{Experiment 1 likelihood by trial, game and opponent faced}\label{fig:exp1-lik-by-tr}
\end{figure}

The same pattern holds for Experiment 2 (Figure
\ref{fig:exp2-lik-by-tr}). Again, the BCH\_BT model with between-game
transfer has the highest likelihood in the early stages of the later
games (apart from stage 1 of the shootout game, where the RL model is
better). In later rounds, the likelihood of the RL model exceeds that of
the BCH model.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp2-lik-by-tr-1}

}

\caption{Experiment2 likelihood by trial, game and opponent faced}\label{fig:exp2-lik-by-tr}
\end{figure}

The fact that the likelihoods of the main strategies considered cross
over in both experiments could be interpreted as indicative that
participants switch between strategies as the games progress. According
to this interpretation, participants base their responses in early
rounds on the learned level of their opponents iterative reasoning,
switching later to learned actions through reinforcement. We use hidden
Markov models to more formally test for strategy switching in
participants' play. In these models, the three strategies (RL, BCH with
between-game transfer, and Nash) correspond to latent states which
determine the overt responses (actions chosen). The models allow for
switching between the states over time, and such switches correspond to
strategy switches. Hidden Markov models assume that an observable action
at time \(t\) depends on a latent state at time \(t\). Second, it is
assumed that the latent state at time \(t\) depends on the latent state
at the previous time \(t-1\). The model is specified by the
state-conditional action distributions (these are provided by the
likelihood of the fitted models), an initial state distribution (the
distribution over the strategies at the initial round), and the
state-transition probabilities (probability of switching from one
state/strategy to another). Initial state probabilities and the
transition probabilities were estimated with the depmixS4 package
(Visser and Speekenbrink 2010). As a statistical test of strategy
switching, we compare the hidden Markov model to a constrained version
which assumes the probability of switching from one strategy to a
different one is 0. This model thus assumes that when players start with
a particular strategy, they continue using it throughout the experiment.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp1-posteriors-plot-1}

}

\caption{Experiment 1 posterior probability of strategies by game and opponent faced}\label{fig:exp1-posteriors-plot}
\end{figure}

In Experiment 1, a likelihood-ratio test shows that the HMM model with
switching fits significantly better than the non-switching one,
\(\chi^2(6) = 211.09\), \(p < .001\). We should note that as the no
switch model involves restricting parameters of the switching model on
the bounds of the parameter space, the asymptotic \(p\)-value of this
test may not be reliable. The switching model (AIC =
\ensuremath{1.463687\times 10^{4}}, BIC =
\ensuremath{1.469256\times 10^{4}}) fits also better than the
non-switching model (AIC = \ensuremath{1.483596\times 10^{4}}, BIC =
\ensuremath{1.484988\times 10^{4}}) according to the AIC and BIC. This
provides further statistical evidence in favour of the hypothesis that
participants switch between strategies. Figure
\ref{fig:exp1-posteriors-plot} depicts the average (across participants)
posterior probabilities of each state (strategy), as a function of trial
and opponent faced. As can be seen, there is evidence of strategy
switching in the FWG and Numbers games: Initially, participants appear
to use a random strategy (in the first round of a game, there is no way
to predict the opponent's action), after which the BCH strategy becomes
dominant. In the later rounds of the games, the RL strategy becomes
dominant, however.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp2-posteriors-plot-1}

}

\caption{Experiment 2 posterior probability of strategies by game and opponent faced}\label{fig:exp2-posteriors-plot}
\end{figure}

In Experiment 2, the switching model (AIC =
\ensuremath{1.687997\times 10^{4}}, BIC =
\ensuremath{1.693681\times 10^{4}}) again fits better than the
restricted non-switching model (AIC =
\ensuremath{1.693841\times 10^{4}}, BIC =
\ensuremath{1.695262\times 10^{4}}), \(\chi^2(6) = 70.44\),
\(p < .001\). The posterior probabilities of the strategies (Figure
\ref{fig:exp2-posteriors-plot}) show very clear evidence of strategy
switching across games and stages, from using a BCH model in the initial
rounds to an RL strategy later on.

\hypertarget{discussion-2}{%
\section{Discussion}\label{discussion-2}}

In this study, we investigated human learning transfer across games by
making human participants play against computer agents with limited
levels of iterated reasoning. We were interested in whether participants
learn about the strategy of their opponent and transfer such knowledge
between games, and whether this is modulated by the similarity between
games and the sophistication of the agent.

The results of our first experiment show that the majority of
participants learnt to adapt to their opponent's strategy over multiple
interactions, and generalised this learning to a similar game.
Performance in early rounds indicated that learning transfer was
moderated by the degree of similarity between games. In the second
experiment, participants faced both types of opponents, allowing for a
stronger test of opponent modelling, as participants would need to learn
a different strategy for each opponent. In Experiment 1, participants
could learn a single strategy for each game, making opponent modelling
possibly less pertinent. In Experiment 2, we again found clear evidence
of transfer in the early rounds of the later games. In addition to the
moderation of transfer by the similarity between games, we also found
that learning transfer was moderated by the type of opponent faced:
Transfer was harder when facing the level-2 opponent as compared to the
level-1 opponent. That evidence for transfer was weaker when facing the
more sophisticated level-2 opponent may be due to a higher difficulty of
learning that opponent's strategy. If it is more difficult to establish
a model of the level-2 opponent, there is likely less knowledge to
transfer to the new game. A player cannot transfer what they have not
learnt.

What exactly did the players learn in the first game (RPS) that allowed
them to beat their opponent in the later games (FWG, and Numbers or
Shootout)? What did the players learn specifically about their
opponent's strategy and what form did this learning take? One possible
answer is that participants learned simple rules based on last round
play. For instance, ``play scissors whenever my opponent played rock in
last round'', or ``play paper whenever the last round play was either
rock or scissors''. These are the type of strategies that are learned by
the model-free reinforcement learning we used in our computational
modelling. While this strategy fitted participants' actions the best
overall, there are at least two reasons why this account is not
satisfactory as a complete description of participants' behaviour.
Firstly, the learned strategies are not transferable to new games. There
is no simple way to map ``play scissors whenever my opponent played rock
in last round'' in the RPS game to ``play grass whenever my opponent
played fire in last round''. Such a mapping may be possible by
translating the rules and structure from RPS to FWG, but model-free
reinforcement learning lacks the tools to do this: Model-free
reinforcement learning would need to start from scratch in each new
game, yet we found evidence that participants could successfully exploit
their opponent's strategy in early rounds of new games. Secondly, a
reinforcement learning strategy would fare equally well against the
level-1 and level-2 opponent. Whilst choosing different actions, the
contingency between the state (last round play) and actions is the same
for both opponents. Yet, we found that participants performed better
against the level-1 opponent compared to the level-2 opponent. The
difference in performance between the two types of opponent indicates
that the actions of the more sophisticated level-2 opponent, or the best
response to these, were somehow more difficult to predict.

We are left with two possible explanations: First, it is possible that
participants discovered a heuristic that allowed them to beat their
opponent without explicitly modelling their strategy, and that this
heuristic is transferable to new games. Because of the cyclicity in
action choices (e.g., rock beats scissors, scissors beats paper, paper
beats rock), it is possible to beat a level-2 opponent most of the time
by following a simple rule: Play in the next round whatever the opponent
played in the last round. This is a rule that wins and is transferable
to other games, as it does not depend on action labels. In the same
vein, a heuristic that beats a level-1 player can be stated as ``Choose
the action that would have been beaten by my previous action''.
Intuitively, it seems that this heuristic for a level-2 player is
simpler than that for a level-1 player, which is at odds with the
finding that participants performed better against the level-1 opponent.

A second explanation is that participants engaged in iterative
reasoning, inferring their opponent's beliefs and countering the
resulting actions. For instance, this would be reasoning of the form
``My opponent expects me to repeat my last action, choosing an action
that would beat my last action. Hence, I will choose the action that
beats their best response'' or ``My opponent thinks I expect them to
repeat their action, hence expecting me to choose the action that beats
their last action. They will therefore choose the action that beats this
one, and hence I should choose the action that beats their best
response.'' Beating a level-1 player, in this account, requires being a
level-2 player, and beating a level-2 player requires being a level-3
player. Intuitively, the additional step of iterative reasoning involved
in beating a level-2 player makes the level-3 strategy more demanding
and difficult to implement, which is consistent with the lower
performance against the level-2 opponent.

The differences in performance between the two players, coupled with the
finding of positive transfer, point to participants engaging in
iterative reasoning, and learning something useful about their
opponent's limitations in this regard. This is the type of learning
encapsulated by our Bayesian Cognitive Hierarchy model. It involves the
evaluation of explicit hypotheses and results in better problem-solving
skills (Mandler 2004). Since it is less context dependent, this type of
learning is generalizable to new situations, akin to the more general
framework of rule-based learning explored by Stahl (Stahl 2000, 2003).

We admit that our implementation in the BCH models does not predict a
performance difference between the types of opponents. Starting with an
equal prior belief over the different levels of sophistication, a BCH
player would perform equally well against the level-1 and level-2
opponent. There are two routes to explain the difference in performance.
Firstly, prior beliefs might be biased against higher-level opponents
(i.e., participants might have believed it is more likely that they
would face level-1 opponent than a level-2 opponent). Secondly, if the
actions of a level-2 opponent are more difficult to predict than those
of a level-1 opponent, this might introduce more noise in the likelihood
of the opponents actions given their level of sophistication. Either of
these mechanisms would explain why learning the strategy of the level-2
opponent is more difficult and slower than learning the strategy of the
level-1 opponent. Using hidden Markov models, we found evidence of
strategy switching between the BCH and RL strategies, and such switching
seems more consistent with the latter idea. If predicting an opponent's
actions through iterative reasoning is cognitively demanding and
error-prone, it is resource-rational to switch to less costly yet
equally successful strategies when these are available (Lieder and
Griffiths 2020). Initially, a model-free reinforcement learning strategy
will be less successful than an iterative reasoning one. However, given
enough experience, it will be on par with an iterative reasoning
strategy. As it involves simple state-action contingencies, a model-free
RL strategy may also be computationally less costly, making it overall
more effective to rely on this than iterative reasoning. This is similar
to the arbitration between model-free and model-based RL (Daw, Niv, and
Dayan 2005; Simon and Daw 2011; Wouter Kool, Joseph T. McGuire, Zev B.
Rosen, Matthew M. Bovinick 2011). In repeated and well-practised
situations, relying on habits allows one to save cognitive resources for
other demands. However, when the environment -- or game -- changes, it
is prudent to use all available resources to reorient oneself.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Our results show that people can successfully deviate from Nash
equilibrium play to exploit deviations from such play by their
opponents. Moreover, people can transfer knowledge about the limitations
of their opponents to new situations. This transfer of a model of the
opponent depends on the similarity between the prior and new game, as
well as the sophistication of the opponent. Transfer is better to
similar games, and for less sophisticated agents. Within games, we found
evidence for a switch from a more reasoning-based strategy which allows
for between-game transfer, to a more habitual strategy which does not.
This is consistent with a rational trade-off between the goal of
maximising performance and minimizing the cost of computing the best
possible actions.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-batzilis}{}%
Batzilis, Dimitris, Sonia Jaffe, Steven Levitt, John A List, and Jeffrey
Picel. 2016. ``How Facebook Can Deepen Our Understanding of Behavior in
Strategic Settings: Evidence from a Million Rock-Paper-Scissors Games.''
working paper.

\leavevmode\hypertarget{ref-camerer2003behavioural}{}%
Camerer, Colin F. 2003. ``Behavioural Studies of Strategic Thinking in
Games.'' \emph{Trends in Cognitive Sciences} 7 (5). Elsevier: 225--31.

\leavevmode\hypertarget{ref-camerer2004cognitive}{}%
Camerer, Colin F, Teck-Hua Ho, and Juin-Kuan Chong. 2004. ``A Cognitive
Hierarchy Model of Games.'' \emph{The Quarterly Journal of Economics}
119 (3). MIT Press: 861--98.

\leavevmode\hypertarget{ref-camerer1997experience}{}%
Camerer, Colin, Teck-Hua Ho, and Others. 1997. ``Experience-weighted
attraction learning in games: A unifying approach.''

\leavevmode\hypertarget{ref-knez2000}{}%
Camerer, Colin, and Marc Knez. 2000. ``Increasing Cooperation in
Prisoner's Dilemmas by Establishing a Precedent of Efficiency in
Coordination Games.''

\leavevmode\hypertarget{ref-cheung1994learning}{}%
Cheung, Yin-Wong, and Daniel Friedman. 1994. \emph{Learning in
evolutionary games: some laboratory results}. University of California,
Santa Cruz.

\leavevmode\hypertarget{ref-costa2001cognition}{}%
Costa-Gomes, Miguel, Vincent P Crawford, and Bruno Broseta. 2001.
``Cognition and Behavior in Normal-Form Games: An Experimental Study.''
\emph{Econometrica} 69 (5). Wiley Online Library: 1193--1235.

\leavevmode\hypertarget{ref-daw2005uncertainty}{}%
Daw, Nathaniel D, Yael Niv, and Peter Dayan. 2005. ``Uncertainty-Based
Competition Between Prefrontal and Dorsolateral Striatal Systems for
Behavioral Control.'' \emph{Nature Neuroscience} 8 (12). Nature
Publishing Group: 1704--11.

\leavevmode\hypertarget{ref-ho2004economics}{}%
Ho, Teck H, Colin F Camerer, and Juin-Kuan Chong. 2004. ``The economics
of learning models: A self-tuning theory of learning in games.''
California Institute of Technology.

\leavevmode\hypertarget{ref-ho1998iterated}{}%
Ho, Teck-Hua, Colin Camerer, and Keith Weigelt. 1998. ``Iterated
dominance and iterated best response in experimental" p-beauty
contests".'' \emph{The American Economic Review} 88 (4). JSTOR: 947--69.

\leavevmode\hypertarget{ref-Lake2017}{}%
Lake, Brenden M., Tomer D. Ullman, Joshua B. Tenenbaum, and Samuel J.
Gershman. 2017. ``Building machines that learn and think like people.''
\emph{Behavioral and Brain Sciences} 40. Cambridge University Press.
\url{https://doi.org/10.1017/S0140525X16001837}.

\leavevmode\hypertarget{ref-lieder2020resource}{}%
Lieder, Falk, and Thomas L Griffiths. 2020. ``Resource-Rational
Analysis: Understanding Human Cognition as the Optimal Use of Limited
Computational Resources.'' \emph{Behavioral and Brain Sciences} 43.
Cambridge University Press.

\leavevmode\hypertarget{ref-mandler2004foundations}{}%
Mandler, Jean Matter. 2004. \emph{The Foundations of Mind: Origins of
Conceptual Thought}. Oxford University Press.

\leavevmode\hypertarget{ref-mertens1990repeated}{}%
Mertens, Jean-François. 1990. ``Repeated games.'' In \emph{Game Theory
and Applications}, 77--130. Elsevier.

\leavevmode\hypertarget{ref-R-DEoptim}{}%
Mullen, Katharine, David Ardia, David Gil, Donald Windover, and James
Cline. 2011. ``DEoptim: An R Package for Global Optimization by
Differential Evolution.'' \emph{Journal of Statistical Software} 40 (6):
1--26. \url{https://doi.org/10.18637/jss.v040.i06}.

\leavevmode\hypertarget{ref-nagel1995unraveling}{}%
Nagel, Rosemarie. 1995. ``Unraveling in Guessing Games: An Experimental
Study.'' \emph{The American Economic Review} 85 (5). JSTOR: 1313--26.

\leavevmode\hypertarget{ref-shachat2004we}{}%
Shachat, Jason, and J Todd Swarthout. 2004. ``Do we detect and exploit
mixed strategy play by opponents?'' \emph{Mathematical Methods of
Operations Research} 59 (3). Springer: 359--73.

\leavevmode\hypertarget{ref-Simon_Daw_11}{}%
Simon, Dylan A., and Nathaniel D. Daw. 2011. ``Environmental statistics
and the trade-off between model-based and TD learning in humans.''
\emph{Advances in Neural Information Processing Systems 24: 25th Annual
Conference on Neural Information Processing Systems 2011, NIPS 2011},
1--9.

\leavevmode\hypertarget{ref-spiliopoulos2013strategic}{}%
Spiliopoulos, Leonidas. 2013. ``Strategic adaptation of humans playing
computer algorithms in a repeated constant-sum game.'' \emph{Autonomous
Agents and Multi-Agent Systems} 27 (1). Springer: 131--60.

\leavevmode\hypertarget{ref-stahl2000rule}{}%
Stahl, Dale O. 2000. ``Rule Learning in Symmetric Normal-Form Games:
Theory and Evidence.'' \emph{Games and Economic Behavior} 32 (1).
Elsevier: 105--38.

\leavevmode\hypertarget{ref-stahl2003sophisticated}{}%
---------. 2003. ``Sophisticated Learning and Learning Sophistication.''
\emph{Available at SSRN 410921}.

\leavevmode\hypertarget{ref-stahl1995players}{}%
Stahl, Dale O, and Paul W Wilson. 1995. ``On players models of other
players: Theory and experimental evidence.'' \emph{Games and Economic
Behavior} 10 (1). Elsevier: 218--54.

\leavevmode\hypertarget{ref-R-depmixS4}{}%
Visser, Ingmar, and Maarten Speekenbrink. 2010. ``depmixS4: An R Package
for Hidden Markov Models.'' \emph{Journal of Statistical Software} 36
(7): 1--21. \url{http://www.jstatsoft.org/v36/i07/}.

\leavevmode\hypertarget{ref-wang2014social}{}%
Wang, Zhijian, Bin Xu, and Hai-Jun Zhou. 2014. ``Social Cycling and
Conditional Responses in the Rock-Paper-Scissors Game.''
\emph{Scientific Reports} 4 (1). Nature Publishing Group: 1--7.

\leavevmode\hypertarget{ref-watkins1992q}{}%
Watkins, Christopher J C H, and Peter Dayan. 1992. ``Q-learning.''
\emph{Machine Learning} 8 (3-4). Springer: 279--92.

\leavevmode\hypertarget{ref-Kool_2011}{}%
Wouter Kool, Joseph T. McGuire, Zev B. Rosen, Matthew M. Bovinick. 2011.
``Decision Making and the Avoidance of Cognitive Demand.''
\emph{Experimental Psychology}.
\url{https://doi.org/10.2996/kmj/1138846322}.

\bibliographystyle{spphys}
\bibliography{Mendeley2.bib}

\end{document}